Performance-portable Parallel Programming in Computer Science Education with Kokkos

### 2020-01-25 src/4.ProgammingModel.tex
Adding other example and back-end example.
Added new picture.

The semantic of the parallel\_for class in Kokkos demonstrates the descriptive nature of the programming model. Programming with patterns and abstractions allows the compiler to generate optimized transformations. 
Generally parallel patterns in Kokkos do not guarantee iteration ordering nor degree of concurrency during their execution. This gives the freedom to the underlying abstraction layers to implement different mapping patterns on different hardware such as assignment of iterations to threads or vector lanes. We believe that these abstraction and their descriptive semantic represent key programming primitives and the way forward for generic support for parallel programming. In conclusion, Figure~\ref{fig:SemCaptureKokkos} shows the semantic capture defined by the Kokkos programming model. 
%Figure~\ref{fig:stack} shows abstraction layers in Kokkos and and how it fits in a parallel programming landscape.
To the interested reader, the complete programming model specification can be accessed on-line\cite{pub:KOKKOS}. Figure~\ref{fig:abstractions} summarizes patterns and abstractions and structures them by information type in the semantic capture.
The next section shows an example of the Kokkos parallel\_for and explains how its mapped to the the underlying back-end programming model.

\begin{figure}
\centerline{\includegraphics[width=0.3\textwidth]{img/Stack.png}}
\caption{Landscape}
\label{fig:stack}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=0.3\textwidth]{img/Abstractions.png}}
\caption{Landscape}
\label{fig:abstractions}
\end{figure}

%\begin{center}
%\begin{tabular}{ |c|c|c| } 
% \hline
% cell1 & cell2 & cell3 \\ 
% cell4 & cell5 & cell6 \\ 
% cell7 & cell8 & cell9 \\ 
% \hline
%\end{tabular}
%\end{center}

%\begin{figure}
%\begin{Verbatim}[frame=leftline]
%double f(){
% typedef Kokkos::Cuda ES;
% typedef Kokkos::CudaSpace MS;
% typedef Kokkos::LayoutLeft LL;
% typedef Kokkos::RangePolicy<ES> Range_policy;
% typedef Kokkos::View<double*, LL, MS> V
% V A( "A", N);
% V::HostMirror h_A = Kokkos::create_mirror_view( A );
% for ( int i = 0; i < N; ++i ) //Initialize on host
%   h_A( i ) = 1;
% Kokkos::deep_copy( A, h_A );
% double result = 0;
% Kokkos::parallel_reduce( "yAx", range_policy( 0, N ), 
%  KOKKOS_LAMBDA ( int i, double &update ) {
%   update += A( i );
% }, result );
% return result;
%}
%\end{Verbatim}
%\caption{Declarative parallel programming provides uses pragma annotations to capture semantic information.}
%\label{figOMPLike}
%\end{figure}

>>>>

The semantic of the parallel\_for class in Kokkos demonstrates the descriptive nature of the programming model. Programming with patterns and abstractions allows the compiler to generate optimized transformations. 
Generally parallel patterns in Kokkos do not guarantee iteration ordering nor degree of concurrency during their execution. This gives the freedom to the underlying abstraction layers to implement different mapping patterns on different hardware such as assignment of iterations to threads or vector lanes. We believe that these abstraction and their descriptive semantic represent key programming primitives and the way forward for generic support for parallel programming. In conclusion, Figure~\ref{fig:SemCaptureKokkos} shows the semantic capture defined by the Kokkos programming model. Figure~\ref{fig:abstractions} summarizes patterns and abstractions and structures them by information type in the semantic capture. To the interested reader, the complete programming model description can be accessed on-line\cite{KOKKOS_WIKI}.
The next section shows an example of the Kokkos parallel\_for and explains how it is mapped to the the underlying back-end programming model.

\begin{figure}
\centerline{\includegraphics[width=0.45\textwidth]{img/Abstractions.png}}
\caption{Overview of abstractions that define the constrained semantic of the Kokkos parallel programming model.}
\label{fig:abstractions}
\end{figure}
